---
title: "Paid Search Advertising Response"
author: "Sandra Hong"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Short-run and Long-run Paid Search Advertising Response  

# Loading Data

```{r, echo = FALSE, message = FALSE}

# Clear All Variables & Clear the Screen
rm(list=ls())
cat("\014")
library(dplyr)
library(lubridate)

# Read in the Data
data.adv = read.csv("~/Desktop/MSBA/Customer_Analytics/Ex3_Advertising/Ex3_Data_R.csv")

# Explore the data
str(data.adv)
summary(data.adv)
```

# Prepare Data
 
In our case, we create new variables: log(Advertising), Advertising^2, and Lag(Sales)

```{r}
# Create log Advertising
data.adv$logTotal.Spent <- log(data.adv$Total.Spent)

# Create Advertising^2
data.adv$Total.Spent2 <- (data.adv$Total.Spent)^2

# Create LAG Sales
data.adv$lagSales <- c(NA, head(data.adv$Visitors, -1))

# Clean date
data.adv$dates <- as.Date(data.adv$Day, "%m/%d/%Y")
data.adv$dates_month <- month(as.POSIXlt(data.adv$dates, format = "%m/%d/%Y"))

# Display the data
summary(data.adv)
```


## Short-run Response.  

To analyze the short-run response of clicks (i.e., visitors) to advertising (i.e., paid search spending), we fit the following three models of advertising-click response. At this stage we will not add other information to the model; keep things simple and stick to advertising.

We created new variables for the Log Total Spent and (Total Spent)^2 to generate the linear, concave logarithmic and concave quadratic regression models. Based on the results, we found that all variables in all the models were < 0.05 and hence, considered significant along with the overall P-values of all models. The R squared value was found to be highest for the concave quadratic model.  

```{r}
#i. Simple linear 
lm.model1 <- lm(Visitors ~ Total.Spent, data = data.adv)
#ii. Concave logarithmic
lm.model2 <- lm(Visitors ~ logTotal.Spent, data = data.adv)
#iii. Concave quadratic 
lm.model3 <- lm(Visitors ~ Total.Spent + Total.Spent2, data = data.adv)

# Display Results
summary(lm.model1)
summary(lm.model2)
summary(lm.model3)
```

We then computed the advertising elasticity implied by each model using July’s monthly clicks and advertising spent to scale the elasticities. The elasticities for linear, logarithmic, and quadratic models are 0.2641779, 0.3013407, and 0.3731448, respectively. 
```{r}

july_calc <- data.adv %>% filter(data.adv$dates_month == 7) %>% summarise(mean_spent=mean(Total.Spent), mean_visitors = mean(Visitors), quotient=mean(Total.Spent) / mean(Visitors))

# coefficients from models
beta_1 = lm.model1$coefficients
beta_2 = lm.model2$coefficients
beta_3 = lm.model3$coefficients

#calculating elasticites 
elasticity_1 = unname(beta_1[2]*july_calc[3])
elasticity_2 = unname(beta_2[2]/july_calc[2])
elasticity_3 = unname((beta_3[2] + 2*beta_3[3]*july_calc[1])*july_calc[3])

#printing elasticities
elasticity_1
elasticity_2 
elasticity_3
```

All our R-squared values were within .08 of each other. All our models also had statistical significance for all coefficients. We are confident that all our models are quite strong and could be used in the short term to project advertising returns. Out of all our models, the concave quadratic model had the highest R squared, so if we were to pick the best model solely based on the highest R square value, that would be the best one. However, a quadratic model is not very realistic in terms of projecting the future, as quadratic models imply future returns that should not be expected in advertising, due to its exponential growth. We believe that the best model is logarithmic, as that has an anti-saturation mechanic built in. Linear is also a fair model, but we believe that the logarithmic model represents the market best. 

## Long-run Response.  
To analyze the long-run response of clicks to advertising using the same data, we it the same three models of advertising response but incorporate an exponentially decaying lag effect for advertising. 

All three models have high adjusted R-square (0.9609, 0.9696 and 0.9603). 

```{r}
#i. Simple linear with carryover
lm.model4 <- lm(Visitors ~ Total.Spent + lagSales, data = data.adv)
#ii. Concave logarithmic with carryover
lm.model5 <- lm(Visitors ~ logTotal.Spent + lagSales, data = data.adv)
#iii. Concave quadratic with carryover
lm.model6 <- lm(Visitors ~ Total.Spent + Total.Spent2 + lagSales, data = data.adv)

# Display Results
summary(lm.model4)
summary(lm.model5)
summary(lm.model6)

```


We then computed the long-run advertising elasticity implied by each model, again using July’s monthly figures for scaling. Elasticities for linear, logarithmic, and quadratic models are 0.3636333, 0.3761668, and 0.383433, respectively. 


```{r}

# coefficients from models
beta_4 = lm.model4$coefficients
beta_5 = lm.model5$coefficients
beta_6 = lm.model6$coefficients

#calculating elasticities
elasticity_4 = unname((beta_4[2]/(1-beta_4[3]))*july_calc[3])
elasticity_5 = unname((beta_5[2]/(1-beta_5[3]))*(1/july_calc[2]))
elasticity_6 = unname(((beta_6[2] + 2*beta_6[3]*july_calc[1])/(1-beta_6[4]))*july_calc[3])

#printing elasticities

elasticity_4 
elasticity_5 
elasticity_6
```


All of our long run models had high and very similar R squared and should be considered acceptable. The goal of the exercise is to find a long-run model with a high correlation of total sales to total advertising spend. Given model 4 has the highest R squared and the lowest elasticity, we would recommend this model be used for future forecasts.
 
# Saturation
Drawing on the models that you fitted above, we computed the saturation level for advertising spending. Results are reported in terms of daily advertising spending. 
 
The saturation level for daily advertising spending in the concave quadratic model without advertising carryover is 1150.32 dollars per day, and the saturation level in the concave quadratic model with advertising carryover is 3447.06 dollars per day.  
 
```{r}

beta = lm.model3$coefficients
sat_lm.model3 = - beta[2] / (2*beta[3])

beta_ca = lm.model6$coefficients
sat_lm.model6 = - beta_ca[2] / (2*beta_ca[3])

#printing saturation
sat_lm.model3
sat_lm.model6
```
```

